---
title: "apollo test"
output: html_document
date: "2023-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# ################################################################# #
#### LOAD LIBRARY AND DEFINE CORE SETTINGS                       ####
# ################################################################# #
```{r}
### Clear memory
rm(list = ls())
setwd("~/Desktop/quanti/QQM-Mobility")

### Load data from package
### if data is to be loaded from a file (e.g. called data.csv), 
### the code would be: database = read.csv("data.csv",header=TRUE)
load("database_PT_final.Rda")

### remove empty rows
library(dplyr)
database_choices=read.csv("finaldata.csv",header=TRUE)

database<-inner_join(database_choices, database, by ='RID')


i<-!is.na(database$pref1)
database<-filter(database,i==TRUE)

database$buy_actually<-ifelse(database$SEQ==1,database$buy_actually_e1,
                       ifelse(database$SEQ==2,database$buy_actually_e2,
                       ifelse(database$SEQ==3,database$buy_actually_e3, 
                       ifelse(database$SEQ==4,database$buy_actually_e4,
                              database$buy_actually_e5 ))))

database$klimaticket<-(database$p_season_1-1)+(database$p_season_2-1)
database$klimaticket<-ifelse(database$klimaticket<2,database$klimaticket,1)

#database<-filter(database,buy_actually==6)
#database<-filter(database,p_mtools_1==2)
#RURAL/LOCATION
#database<-filter(database,p_place_inhab<3)
#PT usage
#database<-filter(database,p_mob_ptintensity<4)
#INCOME
#database<-filter(database,income<4)
#intention to sell car
#database<-filter(database,p_mtools_1==2&buy_sellcar>0)
#travel in rural areas
#database<-filter(database,p_mob_rural==1)
#travel during night time
#database<-filter(database,p_mob_time<3)
#vollzeit
#database<-filter(database,status==1)
#low climate concern (<4)
#database<-filter(database,Statements_1<4)
#has matura
#database<-filter(database,p_matura==1)
#went through survey fast
#database<-filter(database,DURATION<500)

#remove those who never consider buying a mobility guarantee
#database<-filter(database,buy_never==0)
#remove those who always opt to buy mobility guarantee
#database<-filter(database,buy_not>0) 

database <- database %>% rename("Spatial_1" = "a1_x2",
                          "Temporal_1" = "a1_x3",
                          "DRTPattern_1" = "a1_x4",
                          "Waitingtime_1" = "a1_x5",
                          "Discount_1" = "a1_x6",
                          "Compensation_1" = "a1_x7",
                          "Price_1" = "a1_x8",
                          "Spatial_2" = "a3_x2",
                          "Temporal_2" = "a2_x3",
                          "DRTPattern_2" = "a2_x4",
                          "Waitingtime_2" = "a2_x5",
                          "Discount_2" = "a2_x6",
                          "Compensation_2" = "a2_x7",
                          "Price_2" = "a2_x8",
                          "CHOICE" = "pref1"
                          )

save(database,file="database_PT_final_choices.Rda")

#Framing-Variable ist "deck" (1=wie bisher, 2=Mindeststandard) und Kaufen ist 
#"buy_actually_e1 - buy_actually_e5" (6=Nein, 1 und 3=Kaufen ja, aber behalte, 
#4 und 5=Kaufen ja und abschaffen <- 1-5 je nach: ein oder mehr als ein Auto im HH)


### Load Apollo library
library(apollo)

### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "MNL_SP",
  modelDescr      = "Simple MNL model on PT guarantee choice",
  indivID         = "RID", 
  outputDirectory = "output"
)

# ################################################################# #
#### LOAD DATA AND APPLY ANY TRANSFORMATIONS                     ####
# ################################################################# #


# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(
              b_spatial  = 0,
              b_temporal  = 0,
              b_drtpattern  = 0,
              b_waiting = 0,
              b_discount  = 0,
              b_compensation  = 0,
              b_price = 0
#              b_temporal17=0,
#              b_temporal24=0,
#              b_comp_improved=0,
#              b_comp_generous=0
)


### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c()

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){
  
  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))
  
  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  
   V[["alt1"]]  = b_spatial  * Spatial_1 + b_temporal * Temporal_1 + b_drtpattern * DRTPattern_1 +
     b_waiting * Waitingtime_1 + b_discount * Discount_1 + b_compensation * Compensation_1 + b_price * Price_1
   V[["alt2"]]  = b_spatial  * Spatial_2 + b_temporal * Temporal_2 + b_drtpattern * DRTPattern_2 +
     b_waiting * Waitingtime_2 + b_discount * Discount_2 + b_compensation * Compensation_2 + b_price * Price_2

   # V[["alt1"]]  = b_price*(b_spatial  * Spatial_1 + b_temporal * Temporal_1 + b_drtpattern * DRTPattern_1 +
   #   b_waiting * Waitingtime_1 + b_discount * Discount_1 + b_compensation * Compensation_1 + Price_1)
   # V[["alt2"]]  = b_price*(b_spatial  * Spatial_2 + b_temporal * Temporal_2 + b_drtpattern * DRTPattern_2 +
   #   b_waiting * Waitingtime_2 + b_discount * Discount_2 + b_compensation * Compensation_2 + Price_2)
   # 
   
#  V[["alt1"]]  = asc_1 + b_spatial  * Spatial_1 + b_temporal17 * (Temporal_1==17) +b_temporal24 * (Temporal_1==24) + b_drtpattern * DRTPattern_1 +
#    b_waiting * Waitingtime_1 + b_discount * Discount_1 + b_comp_improved * (Compensation_1==2) + b_comp_generous * (Compensation_1==3) + b_price * Price_1
#  V[["alt2"]]  = b_spatial  * Spatial_2 + + b_temporal17 * (Temporal_2==17) +b_temporal24 * (Temporal_2==24) + + b_drtpattern * DRTPattern_2 +
#    b_waiting * Waitingtime_2 + b_discount * Discount_2 + b_comp_improved * (Compensation_2==2) + b_comp_generous * (Compensation_2==3) + b_price * Price_2
  

   
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives  = c(alt1=1, alt2=2), 
    avail         = 1, 
    choiceVar     = CHOICE,
    utilities     = V
  )
  
  ### Compute probabilities using MNL model
  P[["model"]] = apollo_mnl(mnl_settings, functionality)
  
  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model)

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

apollo_saveOutput(model)



deltaMethod_settings=list(expression=c(WTP_spatial="b_spatial/b_price",
                                       WTP_temporal  = "b_temporal/b_price",
                                       WTP_drtpattern  = "b_drtpattern/b_price",
                                       WTP_waiting = "b_waiting/b_price",
                                       WTP_discount  = "b_discount/b_price",
                                       WTP_compensation  = "b_compensation/b_price"))
 
apollo_deltaMethod(model, deltaMethod_settings)


```

### ESTIMATE LATENT CLASS MODEL

# ################################################################# #
# ################################################################# #
####### LATENT CLASS MODEL #############
# ################################################################# #
# ################################################################# #

### Initialise code
apollo_initialise()


apollo_control = list(
  modelName  ="LC_PT_guarantee",
  modelDescr ="LC model",
  indivID    ="RID"
)



# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(b_spatial_1  = 0,
              b_temporal_1  = 0,
              b_drtpattern_1  = 0,
              b_waiting_1 = 0,
              b_discount_1  = 0,
              b_compensation_1  = 0,
              b_price_1 = 0,
              b_spatial_2  = 0,
              b_temporal_2  = 0,
              b_drtpattern_2  = 0,
              b_waiting_2 = 0,
              b_discount_2  = 0,
              b_compensation_2  = 0,
              b_price_2 = 0,
              delta_1 =0,
              delta_2 =1
 #             gamma_framing=0,
 #            gamma_notbought=0
              )

#gamma additional as class membership equations

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("delta_1")

# fix all for b to zero

# ################################################################# #
#### DEFINE LATENT CLASS COMPONENTS                              ####
# ################################################################# #


apollo_lcPars=function(apollo_beta, apollo_inputs){
  lcpars = list()
  lcpars[["b_spatial"]] = list(b_spatial_1, b_spatial_2)
  lcpars[["b_temporal"]] = list(b_temporal_1, b_temporal_2)
  lcpars[["b_drtpattern"]] = list(b_drtpattern_1, b_drtpattern_2)
  lcpars[["b_waiting"]] = list(b_waiting_1,  b_waiting_2)
  lcpars[["b_discount"]] = list(b_discount_1, b_discount_2)
  lcpars[["b_compensation"]] = list(b_compensation_1, b_compensation_2)
  lcpars[["b_price"]] = list(b_price_1, b_price_2)

  
  V=list() #class membership equations
  V[["class_1"]] = delta_1 
  V[["class_2"]] = delta_2 #+ gamma_notbought*buy_never+gamma_framing*deck
  
  mnl_settings = list(
    alternatives = c(class_1=1, class_2=2), 
    avail        = 1, 
    choiceVar    = NA, 
    V            = V
  )
  
  
  lcpars[["pi_values"]] = apollo_mnl(mnl_settings, functionality="raw")
  
  lcpars[["pi_values"]] = apollo_firstRow(lcpars[["pi_values"]], apollo_inputs)
  
  return(lcpars)
}



# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){
  
  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))
  
  ### Create list of probabilities P
  P = list()
  
  
  ### Define settings for MNL model component that are generic across classes
  mnl_settings = list(
    alternatives  = c(alt1=1, alt2=2),
    avail        = list(1),
    choiceVar    = CHOICE
  )
  

  ### Loop over classes
  s=1
  while(s<=2){
    
    ### Compute class-specific utilities
    V = list()
    
    V[["alt1"]]  = b_spatial[[s]]  * Spatial_1 + b_temporal[[s]] * Temporal_1 + b_drtpattern[[s]] * DRTPattern_1 +
      b_waiting[[s]] * Waitingtime_1 + b_discount[[s]] * Discount_1 + b_compensation[[s]] * Compensation_1 + b_price[[s]] * Price_1
    V[["alt2"]]  = b_spatial[[s]]  * Spatial_2 + b_temporal[[s]] * Temporal_2 + b_drtpattern[[s]] * DRTPattern_2 +
      b_waiting[[s]] * Waitingtime_2 + b_discount[[s]] * Discount_2 + b_compensation[[s]] * Compensation_2 + b_price[[s]] * Price_2

    
    mnl_settings$V = V
    mnl_settings$componentName = paste0("Class_",s)
    
    ### Compute within-class choice probabilities using MNL model
    P[[paste0("Class_",s)]] = apollo_mnl(mnl_settings, functionality)
    
    ### Take product across observation for same individual
    P[[paste0("Class_",s)]] = apollo_panelProd(P[[paste0("Class_",s)]], apollo_inputs ,functionality)
    
    s=s+1}
  
  ### Compute latent class model probabilities
  lc_settings   = list(inClassProb = P, classProb=pi_values)
  P[["model"]] = apollo_lc(lc_settings, apollo_inputs, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}
# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

#apollo_beta=apollo_searchStart(apollo_beta, apollo_fixed,apollo_probabilities, apollo_inputs)
#apollo_outOfSample(apollo_beta, apollo_fixed,apollo_probabilities, apollo_inputs)

### Estimate model
model = apollo_estimate(apollo_beta, apollo_fixed, 
                        apollo_probabilities, apollo_inputs,
                        estimate_settings=list(writeIter=FALSE))

### Show output in screen
apollo_modelOutput(model)

### Save output to file(s)
apollo_saveOutput(model)


deltaMethod_settings=list(expression=c(WTP_spatial1="b_spatial_1/b_price_1",
                                       WTP_temporal1  = "b_temporal_1/b_price_1",
                                       WTP_drtpattern1  = "b_drtpattern_1/b_price_1",
                                       WTP_waiting1 = "b_waiting_1/b_price_1",
                                       WTP_discount1  = "b_discount_1/b_price_1",
                                       WTP_compensation1  = "b_compensation_1/b_price_1",
                                       WTP_spatial2="b_spatial_2/b_price_2",
                                       WTP_temporal2  = "b_temporal_2/b_price_2",
                                       WTP_drtpattern2  = "b_drtpattern_2/b_price_2",
                                       WTP_waiting2 = "b_waiting_2/b_price_2",
                                       WTP_discount2  = "b_discount_2/b_price_2",
                                       WTP_compensation2  = "b_compensation_2/b_price_2"))

apollo_deltaMethod(model, deltaMethod_settings)




